Untitled

### 2019-11-12 Paper_resubmit.Rmd
forgot to save before last commit...

```{r, echo=F}
p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))

patient.file = '~/Data/Ellie/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(patient.file)[8:14]
all.val.info = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(patient.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status|Endoscopy|Block|Path'), `Sample Type`, `SLX-ID`, `Index Sequence`,Cohort) %>% 
    dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`))
}))

pastefun<-function(x) {
  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
  return(x)
}

all.val.info = all.val.info %>% rowwise %>% 
  mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup %>%
  #mutate(`Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', '')) %>% 
  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`)

all.val.info$Block = sapply(strsplit(all.val.info$Block, ''), function(x) {
  if (is.na(x)) return(1)
  if (length(x) == 1) {
    if (!grepl('\\d', x)) return(recode(x, 'A'=1, 'B'=2, 'C'=3, 'D'=4, 'E'=5, 'F'=6,'G'=7 ))
    return(as.numeric(x))
  }
  if (length(x) > 1) {
    i = which(grepl('\\d', x))
    return(as.numeric(x[i]))
  }
})

all.val.info = left_join(all.val.info, p.diag, by=c('Hospital Research ID', 'PathID'='Block ID')) %>% group_by(`Hospital Research ID`) %>% 
  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))

val.info = all.val.info %>% filter(Pathology != 'OAC')

@@

all.val.preds = do.call(bind_rows, purrr::map(list.files(vdir, 'predictions.tsv', full.names = T, recursive = T), function(f) {
    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`)
    #left_join(pd, all.val.info, by=c('Hospital Research ID', 'Sample'='Samplename', 'Block ID')) 
    return(pd)
})) %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T)) %>%
  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants')

all.val.preds = left_join(all.val.preds, all.val.info, by=c('Hospital Research ID', 'Sample'='Samplename'))

val.preds = all.val.preds %>% dplyr::filter(Sample %in% val.info$Samplename)

oac.preds = all.val.preds %>% filter(Pathology == 'OAC')
#all.val.preds %>% filter(Cohort == 'Old progressor scrolls') %>% group_by(Status, Risk) %>% tally
#val.info %>% filter(Cohort != 'Old progressor scrolls' & Pathology != 'OAC') %>% group_by(Status) %>% tally
```

@@

val.preds = val.preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','GM','ID','LGD'), ordered=T)) %>% 
  dplyr::rename(Samplename = 'Sample') %>% 
  mutate(Final.Endoscopy = `Diagnostic Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))

rp = plot.risk.by.path(val.preds) + labs(title='Validation samples predicted by pathology')  
rp
ggsave(filename='plots/val_pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)

@@

val.preds = do.call(bind_rows, lapply(unique(val.preds$`Hospital Research ID`) %>% 
                            map( function(p) val.preds %>% 
                                   filter(`Hospital Research ID` == p) %>% 
                                   mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
                                 ), transform.by.time, info=val.info) )

@@

for (pt in unique(v.prog$`Hospital Research ID`)) {

  patient = v.prog %>% filter(`Hospital Research ID` == pt) %>% 
    dplyr::select(`Hospital Research ID`,Status, Endoscopy, months.before.final, Pathology,Block,Risk) %>% 
    dplyr::rename(GEJ.Distance = Block, Patient = 'Hospital Research ID')

    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)

    return(p)
}

>>>>

```{r, echo=F}
#p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
#  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
pastefun<-function(x) {
  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
  return(x)
}

patient.file = '~/Data/Ellie/QDNAseq/validation/sWGS_validation_batches.xlsx'

oac.info = readxl::read_xlsx(patient.file, 'OAC Test Samples') %>% 
  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
  dplyr::select(-matches('SLX|Index')) %>%
  mutate(Status = factor(Status, levels=levels(pg.samp$Status)), Block = factor(Block, levels=c('E','1','2'), ordered=T))

val.info = readxl::read_xlsx(patient.file, 'Final Validation Samples') %>% 
  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`) %>%
  dplyr::select(-matches('SLX|Index'), -`Path Notes`) %>%
  mutate(Status = factor(Status, levels=levels(pg.samp$Status))) %>% ungroup %>% group_by(`Hospital Research ID`) %>% 
  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Last Endoscopy`, `Endoscopy`, units='weeks')))

val.info$Block = sapply(strsplit(val.info$Block, ''), function(x) {
  if (is.na(x)) return(1)
  if (length(x) == 1) {
    if (!grepl('\\d', x)) return(recode(x, 'A'=1, 'B'=2, 'C'=3, 'D'=4, 'E'=5, 'F'=6,'G'=7 ))
    return(as.numeric(x))
  }
  if (length(x) > 1) {
    i = which(grepl('\\d', x))
    return(as.numeric(x[i]))
  }
})

@@

residuals = do.call(bind_rows, purrr::map(list.files( dirname(vdir), 'residuals.tsv', full.names = T, recursive = T ), function(f) {
  read_tsv(f, col_types = 'ccddddl')
}))

failedQC = residuals %>% filter(varMAD_median > 0.008)

all.val.preds = do.call(bind_rows, purrr::map(list.files(vdir, 'predictions.tsv', full.names = T, recursive = T), function(f) {
    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`)
    return(pd)
})) %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T)) %>%
  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants')

oac.preds = all.val.preds %>% filter(`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`))

all.val.preds = all.val.preds %>% filter( !Sample %in% failedQC$samplename &  
                                            !`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`) )

val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% dplyr::select(Patient, everything())

val.preds = val.preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','GM','ID','LGD'), ordered=T)) %>% 
  mutate(Final.Endoscopy = `Last Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))

@@

```{r}
rp = plot.risk.by.path(val.preds) + labs(title='Validation samples predicted by pathology')  
rp
ggsave(filename='plots/val_pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)

@@

val.preds = dplyr::rename(val.preds, 'Samplename'='Sample')

val.preds = do.call(bind_rows, lapply(unique(val.preds$Patient) %>% 
                            map( function(p) val.preds %>% filter(Patient == p) %>% 
                                   mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
                                 ), transform.by.time, info=val.info) )

for (pt in unique(val.preds$Patient)) {
print(pt)
  patient = val.preds %>% filter(Patient == pt) %>% 
    dplyr::select(Patient, Status, Endoscopy, months.before.final, Pathology,Block,Risk) %>% 
    dplyr::rename(GEJ.Distance = Block)

    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F) 
}

@@
